---
title: "Covid Project Group 3 Exploration 3: Counts and Rates by Continent"
author: "Kaitlyn Won (Data Processing for Rates/Visualization), Jyoti Michael D Rozario (Data Processing for Counts), BaoLong Nguyen (Data Analysis)"
output:
  html_document:
    number_sections: yes
    toc: yes  
  pdf_document:
    toc: yes
---
```{r setup, warning=FALSE}
```

# Data Processing/Visualization: Total Confirmed Cases by Continent
Read in all the csv files per continent.
```{r}
africa = read.csv("refined_data_Africa.csv")
asia = read.csv("refined_data_Asia.csv")
australia = read.csv("refined_data_Australia.csv")
europe = read.csv("refined_data_Europe.csv")
north_america = read.csv("refined_data_North_America.csv")
south_america = read.csv("refined_data_South_America.csv")
```

Total number of confirmed cases per continent calculated 
```{r}
africa_total_cases = sum(africa$Total.Confirmed.Cases)
asia_total_cases = sum(asia$Total.Confirmed.Cases)
australia_total_cases = sum(australia$Total.Confirmed.Cases)
europe_total_cases = sum(europe$Total.Confirmed.Cases)
na_total_cases = sum(north_america$Total.Confirmed.Cases)
sa_total_cases = sum(south_america$Total.Confirmed.Cases)
total_cases_sum = africa_total_cases + asia_total_cases + australia_total_cases + europe_total_cases + na_total_cases + sa_total_cases
```

Plotted the relative frequencies of the total counts of the confirmed cases
```{r}
total_cases = c(africa_total_cases/total_cases_sum, asia_total_cases/total_cases_sum, australia_total_cases/total_cases_sum, europe_total_cases/total_cases_sum, na_total_cases/total_cases_sum, sa_total_cases/total_cases_sum)
continents = c("Africa", "Asia", "Australia", "Europe", "NA", "SA")
barplot(total_cases, main="Relative Frequency of Total Confirmed Cases by Continent",names.arg=continents, ylim=c(0,0.3))
```

# Data Analysis: Total Confirmed Cases by Continent

# Data Processing/Visualization: Covid Rate by Continent

Installed/added necessary packages and csv file. Notably, the countrycode package is necessary to map countries to continents since this information isn't included in the time series csv. 
```{r}
library(ggplot2)
library(dplyr)
library(remotes) # do install.packages("remotes") if error
# below comment is for countrycode package installation 
# install_github('vincentarelbundock/countrycode') 
library(countrycode)

ts = read.csv("time_series_covid19_confirmed_global.csv")
```

Grouped data by continent (note that the countrycode package groups North and South America as the Americas) and sum each day's covid confirmation case counts per country. Plotted these counts. 
```{r}
ts$Continent = countrycode(ts$Country.Region, "country.name", "continent")

# Had to manually add continents for the few that did not map. Diamond Princess is not a country; it is a 
# cruise ship that was docked in Yokohama, Japan when the coronavirus outbreak happened within the vessel.
# Similarly, MS Zaandam is a cruise ship that was near Panama City when the outbreak happened. 
# Kosovo is a country in Europe.
ts$Continent[which(ts$Country.Region == "Diamond Princess")] = "Asia"
ts$Continent[which(ts$Country.Region == "MS Zaandam")] = "Americas"
ts$Continent[which(ts$Country.Region == "Kosovo")] = "Europe"

ts = select(ts, -c(Province.State, Country.Region, Lat, Long))

ts_by_conts = ts %>% group_by(Continent) %>% summarise_each(funs(sum))
```
Time series all split by continent in separate variables
```{r}
africa_ts = ts_by_conts[ts_by_conts$Continent == "Africa",]
americas_ts = ts_by_conts[ts_by_conts$Continent == "Americas",]
asia_ts = ts_by_conts[ts_by_conts$Continent == "Asia",]
europe_ts = ts_by_conts[ts_by_conts$Continent == "Europe",]
oceania_ts = ts_by_conts[ts_by_conts$Continent == "Oceania",]
```

Function to plot time series by continent 
```{r}
library(rlist)

create_ts_plot <- function(ts) {
  dates = seq(as.Date("2020/1/22"), by="day", length.out = ncol(ts) - 1)
  counts = rep(0, ncol(ts) - 1)
  c = 0
  for (i in colnames(ts)) {
    counts[c] = as.integer(ts[[paste(i)]])
    c = c + 1
  }
  df = data.frame(date = dates, count = counts)
  ggplot(data = df, aes(date, count)) + geom_line()
}
```

Use create_ts_plot function plot time series for each continent's confirmed covid cases
```{r}
create_ts_plot(africa_ts)
create_ts_plot(americas_ts)
create_ts_plot(asia_ts)
create_ts_plot(europe_ts)
create_ts_plot(oceania_ts)
```

# Data Analysis: Covid Rate by Continent



